---
title: "project"
author: "Utkarsh Upadhyay"
date: "2023-04-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("effsize")
library("readxl")
library("car")
library("rstatix")
library("ggplot2")
```

```{r}
ds_file <- "dataset.csv"
data <- read.csv(ds_file)
```
# Hypothesis 1

```{r}
data$grades <- data$Curricular.units.1st.sem..grade. + data$Curricular.units.2nd.sem..grade.
grades_wos <- data$grades[data$Scholarship.holder == 0]
grades_ws <- data$grades[data$Scholarship.holder == 1]

qqnorm(grades_ws, main="QQplot for Grades With Scholarship")
qqline(grades_ws)

shapiro.test(grades_ws)
```

```{r}
qqnorm(grades_wos, main="QQplot for Grades Without Scholarship")
qqline(grades_wos)

shapiro.test(grades_wos)
```

```{r}
data$Scholarship.holder <- as.factor(data$Scholarship.holder)
ggplot(data, aes(x = grades, fill = Scholarship.holder)) +
  geom_density(alpha = 0.5) +
  theme_bw() +
  labs(
    title = "Density plot for Grades of Scholarship holders and non-holders",
    x = "Grades",
    y = "Density"
  ) +
  geom_vline(aes(xintercept = mean(grades_wos, na.rm = TRUE)), color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = mean(grades_ws, na.rm = TRUE)), color = "blue", linetype = "dashed", size = 1)
```

```{r}
print("Mean grade of scholarship holders: ")
mean(grades_ws)
print("Mean grade of non-scholarship holders: ")
mean(grades_wos)
wilcox.test(grades_ws, grades_wos, alternative = "g")
# 
cohen.d(grades_ws, grades_wos)
```
## Conculsion

We reject null hypothesis that scholarship holders does not get greater marks than students with no scholarship.


# Hypothesis 2

## Course vs Tuition fee 
```{r}
course <- data$Course
tuition <- data$Tuition.fees.up.to.date

contigency.table <- table(tuition, course)
print(contigency.table)

expected <- outer(
  rowSums(contigency.table),
  colSums(contigency.table)
) / sum(contigency.table)
rownames(expected) <- c("No tuition", "Tuition")
round(expected, 2)

result <- chisq.test(contigency.table, simulate.p.value = TRUE)

print(result)

# effect size
effect.size <- sqrt(result$statistic / (sum(contigency.table)))
print(effect.size)

residuals <- contigency.table - expected
rownames(residuals) <- c("No tuition", "Tuition")
residuals
colSums(expected)
residuals <- round(t(t(residuals) / colSums(expected)), 2)
residuals

plot(residuals)
```

## Grades vs Tuition fee
  
```{r}
grades_w_fee <- data$grades[data$Tuition.fees.up.to.date == 1]
grades_wo_fee <- data$grades[data$Tuition.fees.up.to.date == 0]

qqnorm(grades_w_fee)
qqline(grades_w_fee)
title("QQplot for Grades With Tuition fee upto date")

shapiro.test(grades_w_fee)

qqnorm(grades_wo_fee)
qqline(grades_wo_fee)
title("QQplot for Grades Without Tuition fee upto date")

shapiro.test(grades_wo_fee)
```

As we can see the data is not normal.
```{r}
data$Tuition.fees.up.to.date <- as.factor(data$Tuition.fees.up.to.date)
ggplot(data, aes(x = grades, fill = Tuition.fees.up.to.date)) +
  geom_density(alpha = 0.5) +
  theme_bw() +
  labs(
    title = "Density plot for Grades of Students with and \
    without tuition fee upto date",
    x = "Grades",
    y = "Density"
  ) +
  geom_vline(aes(xintercept = mean(grades_wo_fee, na.rm = TRUE)),
    color = "red", linetype = "dashed", size = 1
  ) +
  geom_vline(aes(xintercept = mean(grades_w_fee, na.rm = TRUE)),
    color = "blue", linetype = "dashed", size = 1
  )
```

```{r}
print("Mean grade of students with tuition fee upto date: ")
mean(grades_w_fee)
print("Mean grade of students without tuition fee upto date: ")
mean(grades_wo_fee)
wilcox.test(grades_w_fee, grades_wo_fee)
cohen.d(grades_w_fee, grades_wo_fee)
```

Since the p-value is less than 0.05, we reject the null hypothesis that students with tuition fee upto date do get different marks than students without tuition fee upto date.

# Hypothesis 3

## Total grades

```{r}
# Covnert sem1 grades to numeric
data$grades <- as.numeric(data$grades)

dropout.grades <- data$grades[data$Target == "Dropout"]
graduate.grades <- data$grades[data$Target == "Graduate"]
enrolled.grades <- data$grades[data$Target == "Enrolled"]

shapiro.test(dropout.grades)
shapiro.test(graduate.grades)
shapiro.test(enrolled.grades)

## Not normal
```

```{r}
# Perform Kruskal-Wallis ANOVA test
kruskal.test(grades ~ Target, data = data)
```

# As the p-value is less than the significance level 0.05, we can conclude that there are significant differences between the Target groups.

```{r}
# Perform Mann-Whitney U post-hoc test
pairwise.wilcox.test(data$grades, data$Target, p.adjust.method = "bonferroni")

# Effect size of Kruskal-Wallis test
kruskal_effsize(data, grades ~ Target)
```

```{r}
# Density-plots for all targets (using ggplot2)
ggplot(data, aes(x = grades, fill = Target)) +
  geom_density(alpha = 0.5) +
  theme_bw() +
  labs(
    title = "Density plot for grades of different targets",
    x = "Grades",
    y = "Density"
  ) +
  geom_vline(aes(xintercept = mean(dropout.grades, na.rm = TRUE)), color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = mean(graduate.grades, na.rm = TRUE)), color = "green", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = mean(enrolled.grades, na.rm = TRUE)), color = "blue", linetype = "dashed", size = 1)
```

## First-sem grades

```{r}
# Covnert sem1 grades to numeric
data$Curricular.units.1st.sem..grade. <- as.numeric(data$Curricular.units.1st.sem..grade.)

dropout.Curricular.units.1st.sem..grade. <- data$Curricular.units.1st.sem..grade.[data$Target == "Dropout"]
graduate.Curricular.units.1st.sem..grade. <- data$Curricular.units.1st.sem..grade.[data$Target == "Graduate"]
enrolled.Curricular.units.1st.sem..grade. <- data$Curricular.units.1st.sem..grade.[data$Target == "Enrolled"]

shapiro.test(dropout.Curricular.units.1st.sem..grade.)
shapiro.test(graduate.Curricular.units.1st.sem..grade.)
shapiro.test(enrolled.Curricular.units.1st.sem..grade.)

## Not normal
```

```{r}
# Perform Kruskal-Wallis ANOVA test
kruskal.test(Curricular.units.1st.sem..grade. ~ Target, data = data)
```

# As the p-value is less than the significance level 0.05, we can conclude that there are significant differences between the Target groups.

```{r}
# Perform Mann-Whitney U post-hoc test
pairwise.wilcox.test(data$Curricular.units.1st.sem..grade., data$Target, p.adjust.method = "bonferroni")

# Effect size of Kruskal-Wallis test
kruskal_effsize(data, Curricular.units.1st.sem..grade. ~ Target)
```

```{r}
# Density-plots for all targets (using ggplot2)
ggplot(data, aes(x = Curricular.units.1st.sem..grade., fill = Target)) +
  geom_density(alpha = 0.5) +
  theme_bw() +
  labs(
    title = "Density plot for Curricular.units.1st.sem..grade. of different targets",
    x = "Curricular.units.1st.sem..grade.",
    y = "Density"
  ) +
  geom_vline(aes(xintercept = mean(dropout.Curricular.units.1st.sem..grade., na.rm = TRUE)), color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = mean(graduate.Curricular.units.1st.sem..grade., na.rm = TRUE)), color = "green", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = mean(enrolled.Curricular.units.1st.sem..grade., na.rm = TRUE)), color = "blue", linetype = "dashed", size = 1)
```

## Second-sem grades

```{r}
# Covnert sem1 grades to numeric
data$Curricular.units.2nd.sem..grade. <- as.numeric(data$Curricular.units.2nd.sem..grade.)

dropout.Curricular.units.2nd.sem..grade. <- data$Curricular.units.2nd.sem..grade.[data$Target == "Dropout"]
graduate.Curricular.units.2nd.sem..grade. <- data$Curricular.units.2nd.sem..grade.[data$Target == "Graduate"]
enrolled.Curricular.units.2nd.sem..grade. <- data$Curricular.units.2nd.sem..grade.[data$Target == "Enrolled"]

shapiro.test(dropout.Curricular.units.2nd.sem..grade.)
shapiro.test(graduate.Curricular.units.2nd.sem..grade.)
shapiro.test(enrolled.Curricular.units.2nd.sem..grade.)

## Not normal
```

```{r}
# Perform Kruskal-Wallis ANOVA test
kruskal.test(Curricular.units.2nd.sem..grade. ~ Target, data = data)
```

As the p-value is less than the significance level 0.05, we can conclude that there are significant differences between the Target groups.

```{r}
# Perform Mann-Whitney U post-hoc test
pairwise.wilcox.test(data$Curricular.units.2nd.sem..grade., data$Target, p.adjust.method = "bonferroni")

# Effect size of Kruskal-Wallis test
kruskal_effsize(data, Curricular.units.2nd.sem..grade. ~ Target)
```

```{r}
# Density-plots for all targets (using ggplot2)
ggplot(data, aes(x = Curricular.units.2nd.sem..grade., fill = Target)) +
  geom_density(alpha = 0.5) +
  theme_bw() +
  labs(
    title = "Density plot for Curricular.units.2nd.sem..grade. of different targets",
    x = "Curricular.units.2nd.sem..grade.",
    y = "Density"
  ) +
  geom_vline(aes(xintercept = mean(dropout.Curricular.units.2nd.sem..grade., na.rm = TRUE)), color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = mean(graduate.Curricular.units.2nd.sem..grade., na.rm = TRUE)), color = "green", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = mean(enrolled.Curricular.units.2nd.sem..grade., na.rm = TRUE)), color = "blue", linetype = "dashed", size = 1)
```

# ::: {.hypothesis #label name="hypothesis name"}

:::

# Hypothesis 4

```{r}
# from data, extract grades for each country in different arrays
# and then perform Shapiro-Wilk test to check if the data is normal

grades_wrt_country <- list()
countries <- unique(data$Nacionality)
countries <- countries[order(countries)]
for (country in countries) {
  print(country)
  grades_wrt_country[[country]] <- data$grades[data$Nacionality == country]
}

# sort the list by length of the array
# grades_wrt_country <- grades_wrt_country[order(sapply(grades_wrt_country, length), decreasing = TRUE)]

lens <- sapply(grades_wrt_country, length)
lens
```
Since some countries have very low number of students, we will not consider countries with less than 5 students.

```{r}
STUDENTS_CUTOFF <- 5

# remove countries with less than 5 students
grades_wrt_country <- grades_wrt_country[lens >= STUDENTS_CUTOFF]
countries <- countries[lens >= STUDENTS_CUTOFF]
data_country <- data[data$Nacionality %in% countries, ]

# perform Shapiro-Wilk test on each country
shapiro.test_results <- lapply(grades_wrt_country, shapiro.test)

# check if the data is normal
is_normal <- lapply(shapiro.test_results, function(x) x$p.value > 0.05)

print("Countries with normal data:")
print(countries[is_normal == TRUE])
print("Countries with non-normal data:")
print(countries[is_normal == FALSE])
```
Since only one country has normal data, normality of data does not hold.
Hence, we will perform Kruskal-Wallis test on all countries.

```{r}
# perform Kruskal-Wallis test on all countries
kruskal.test_results <- kruskal.test(grades ~ Nacionality, data = data_country)
kruskal.test_results
```

```{r}
# print means of each countries grades
aggregate(grades ~ Nacionality, data = data_country, mean)

# Density-plots for all countries in the array countries
# with mean vertical lines
dd <- data_country
dd$Nacionality <- as.factor(dd$Nacionality)
ggplot(dd, aes(x = grades, fill = Nacionality)) +
  geom_density(alpha = 0.5) +
  theme_bw() +
  labs(
    title = "Density plot for grades of different countries",
    x = "grades",
    y = "Density"
  ) +
  geom_vline(aes(xintercept = mean(grades[data_country$Nacionality == 1],
    na.rm = TRUE
  )), color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = mean(grades[data_country$Nacionality == 3],
    na.rm = TRUE
  )), color = "yellow", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = mean(grades[data_country$Nacionality == 9],
    na.rm = TRUE
  )), color = "green", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = mean(grades[data_country$Nacionality == 10],
    na.rm = TRUE
  )), color = "blue", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = mean(grades[data_country$Nacionality == 12],
    na.rm = TRUE
  )), color = "#433bb2", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = mean(grades[data_country$Nacionality == 14],
    na.rm = TRUE
  )), color = "purple", linetype = "dashed", size = 1)
```

# Hypothesis 5

```{r}
debtors_age <- data$Age.at.enrollment[data$Debtor == 1]
non_debtors_age <- data$Age.at.enrollment[data$Debtor == 0]
all_ages <- data$Age.at.enrollment

print("Lengths of debtors, non-debtors and all ages:")
print(length(all_ages))
print(length(debtors_age))
print(length(non_debtors_age))

print("Means of debtors, non-debtors and all ages:")
mean(debtors_age)
mean(non_debtors_age)
mean(all_ages)

print("Performing normality test:")
shapiro.test(debtors_age)
shapiro.test(non_debtors_age)
shapiro.test(all_ages)

```
Since, the two categories with significantly different sample sizes, we would perform independent-samples randomization test.
```{r}
permsum <- replicate(10000, mean(sample(all_ages, length(debtors_age))))
permsum <- c(permsum, mean(debtors_age))
p_value <- sum(permsum >= mean(debtors_age)) / length(permsum)

hist(permsum, )
abline(v = mean(debtors_age), col = "red")
abline(v = mean(non_debtors_age), col = "green")
# add legend for ablines
legend(
  "topright",
  legend = c("Mean of debtors", "Mean of non-debtors"),
  col = c("red", "green"),
  lty = 1,
  cex = 0.8
)
p_value
```
Since, the p-value is less than 0.05, we reject the null hypothesis that debtors have lower age than non-debtors.

```{r}
# box plot of age of debtors and non-debtors
data$Debtor <- as.factor(data$Debtor)
ggplot(data, aes(x = Debtor, y = Age.at.enrollment)) +
  geom_boxplot() +
  theme_bw() +
  labs(
    title = "Box plot of age of debtors and non-debtors",
    x = "Debtor",
    y = "Age at enrollment"
  )
```

```{r}
```



```{r}
cohen.d(debtors_age, non_debtors_age)
```
We have small effect size.
```{r}
```

```{r}

```
